Opt-K

### 2013-03-15 paper/paper.tex
Added change to likelihood.py to allow you to compute Liz Kellogg's (Baker Lab) likelihood function. Also added stuff to the paper.tex as well as likelihoods.ipynb

$$ 
P[x_{0...T-1}] dx^N = \prod_{i=0}^{T-1} T(x_i \rightarrow x_{i+1}) \cdot \prod_{i=0}^{T} p(x_{i} | \sigma(x_{i}))
$$

@@

$$ 
P[x_{0...T-1}] dx^N = \prod_{i=0}^{T-1} T(x_i \rightarrow x_{i+1}) \cdot \prod_{i=0}^T \frac{1}{V_{\sigma(x_{i})}}
$$

@@

Briefly, we generated uniform random vectors within a bounding box whose size was slightly larger than the data. This random sample was then accepted if it was 'close enough' to a generator in the 100 state model, where close enough meant it was within 0.2 units from any generator in that model. This was used to approximate the convex hull of the data. Then the accepted points were assigned to a state. Then the volume of a state is proportional to the number of samples assigned to that state divided by the total number of samples. We note, that these volumes are actually in units of the total acceptable volume. This affects the absolute value of the likelihood, however, since the total volume was constant in all cases, the relative likelihoods are the same.

\begin{figure}[h!]
\includegraphics[width=3in]{figs/muller_pot_traj.pdf}
\caption{The M\"{u}ller potential consists of three metastable wells. We simulated Langevin dynamics on this potential as validation of the new likelihood function.}
\label{fig:muller_pot}
\end{figure}

\begin{figure}
\includegraphics[width=2in]{figs/like_comp.pdf}
\end{figure}

>>>>

\begin{equation}
\label{eq:like}
P[x_{0...T-1}] dx^N = \prod_{i=0}^{T-1} T(x_i \rightarrow x_{i+1}) \cdot \prod_{i=0}^{T} p(x_{i} | \sigma(x_{i}))
\end{equation}

@@

\begin{equation}
\label{eq:like_vol}
P[x_{0...T-1}] dx^N = \prod_{i=0}^{T-1} T(x_i \rightarrow x_{i+1}) \cdot \prod_{i=0}^T \frac{1}{V_{\sigma(x_{i})}}
\end{equation}

@@

Briefly, we generated uniform random vectors within a bounding box whose size was slightly larger than the data. This random sample was then accepted if it was 'close enough' to a generator in the 100 state model, where close enough meant it was within 0.2 units from any generator in that model. This was used to approximate the convex hull of the data. Then the accepted points were assigned to a state. Then the volume of a state is proportional to the number of samples assigned to that state divided by the total number of samples. We note, that these volumes are actually in units of the total acceptable volume. This affects the absolute value of the likelihood, however, since the total volume was constant in all cases, the relative likelihoods are the same.
\begin{figure}[h!]
\includegraphics[width=3in]{figs/muller_pot_traj.pdf}
\caption{The M\"{u}ller potential consists of three metastable wells. We simulated Langevin dynamics on this potential as a validation of the new likelihood function.}
\label{fig:muller_pot}
\end{figure}

We built an MSM using a fixed lag time for each state decomposition and then calculated the likelihood according to \cref{eq:like_vol}. These likelihoods increased rapidly with $k$, but then plateaued at a value of around 300 states \cref{fig:like_kcenters}. At this plateau, we can conclude that we do not gain any new information by adding new states to our model and so should pick $k$ to be this point. This is a heuristic argument that can be made more mathematically rigorous with the introduction of functions like the Bayesian Information Content (BIC), which is defined in \cref{eq:bic}.

\begin{equation}
BIC = -2 \log L + m \log(n) 
\label{eq:bic}
\end{equation} Here, $L$ denotes the likelihood of a model, while $m$ is the number of parameters used in the model, and $n$ is the number of data points. Essentially this is a mathematical way of saying we would like to use a model with the maximum likelihood, but without introducing too many parameters. 
% don't know how to say this yet:
%We note that we cannot take this too seriously, and there are many other such criteria, such as the Aikaike Information Content (AIC). 
We found that the BIC of the models built using K-Centers clustering has a minimum at $k=250$. This minimum is the ``optimal'' model according to the BIC of our likelihood. As a check of the usefulness of this minimum, we plotted the three slowest implied timescales in the models as a function of $k$. The optimal model was close to the point at which the implied timescales became constant.

\begin{figure}
\includegraphics[width=2in]{figs/like_comp.pdf}
\caption{For models built between 50 and 1,000 states using the K-Centers algorithm, the log likelihood function increased quickly and plateaued at approximately 300 states. Heuristically, at this point, adding another state does not add anything to the model. As a result, by using a criterion, like the BIC, that penalizes using additional parameters to fit the data, we can find an ``optimal'' model. Interestingly, the BIC optimal model occurs at 250 states, which also corresponds to where the three slowest timescales plateau.}
\label{fig:like_kcenters}
\end{figure}
